---
title: "Why put crimes on maps?"
output:
  xaringan::moon_reader:
    css: ["ucl-basic.css", "animate.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

# set plots to have a transparent background, see
# https://gist.github.com/cboettig/5600558
knitr::opts_chunk$set(cache = TRUE, dev.args = list(bg = "transparent"))
```

```{r load packages, include=FALSE}
library("ggrepel")
library("sf")
library("tidyverse")
```



class: speaker-notes

"In this video I'm going to talk about why maps are useful for understanding crime"

"Crime is a geographic phenomenon – all types of crime are concentrated in particular places, and understanding the relationships between crime and places can be really useful for crime prevention"

---

class: green-screen, animated, fadeIn

<!-- blank slide for green screen fade -->



---

class: green-screen, big-picture, animated, fadeIn

```{r load violence data, include=FALSE}
# load crime counts
ew_violence_in_grid <- read_rds(here::here("data/ew_violence_in_grid.Rds"))

# load police force areas
forces <- st_read(
  "https://opendata.arcgis.com/datasets/1fc3e34edece4cd58dabac422c3556c9_0.geojson",
  quiet = TRUE
) %>%
  janitor::clean_names() %>%
  select(pfa19cd, pfa19nm) %>%
  # convert to British National Grid to match the crimes layer
  st_transform(27700)

# local authority districts
districts <- st_read(
  "https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson",
  quiet = TRUE
) %>%
  janitor::clean_names() %>%
  # filter out all except England and Wales
  filter(str_detect(lad19cd, "^E") | str_detect(lad19cd, "^W")) %>%
  select(lad19cd, district = lad19nm) %>%
  # convert to British National Grid to match the crimes layer
  st_transform(27700)
```

```{r create outlines, include=FALSE}
# create outline of England and Wales
ew_outline <- forces %>% st_union() %>% st_as_sf()

# create outline of London
london_outline <- forces %>% 
  filter(pfa19nm %in% c("London, City of", "Metropolitan Police")) %>% 
  st_union() %>% 
  st_as_sf() %>% 
  mutate(region = "London")
```

```{r national violence map, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=13}
# create labels for major cities
ew_labels <- tribble(
  ~label, ~x, ~y,
  "London", -0.1278, 51.5074,
  "Birmingham", -1.8904, 52.4862,
  "Manchester", -2.2426, 53.4808,
  "Newcastle", -1.6178, 54.9783,
  "Bristol", -2.5879, 51.4545
) %>% 
  st_as_sf(coords = c("x", "y"), remove = FALSE, crs = 4326) %>% 
  st_transform(27700) %>% 
  mutate(
    x = unname(st_coordinates(.)[, 1]),
    y = unname(st_coordinates(.)[, 2])
  )

# plot map
ew_violence_in_grid %>%
  mutate(
    group = ifelse(cperc < 0.5, "half of all\nviolence", NA_character_)
  ) %>%
  filter(!is.na(group)) %>%
  ggplot() +
  geom_sf(data = ew_outline, colour = "white", fill = "white", size = 4) +
  geom_sf(data = ew_outline, colour = NA, fill = "white") +
  geom_raster(aes(x = x, y = y, fill = group), na.rm = TRUE, fill = "#F6BE00") +
  geom_sf(data = forces, colour = "grey40", fill = NA) +
  ggrepel::geom_label_repel(
    aes(x = x, y = y, label = label), data = ew_labels, colour = "grey20", 
    label.size = NA, min.segment.length = 0
  ) +
  geom_sf(data = ew_labels) +
  # constrain plot height to area between Portland Bill and Newcastle upon Tyne
  lims(x = range(ew_violence_in_grid$x), y = c(70167, 565390)) + 
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = margin(0, 0, 0, 0)
  )
```



---

class: green-screen, big-picture, animated, fadeIn

```{r london violence map, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
# create labels for major landmarks
london_labels <- tribble(
  ~label, ~x, ~y,
  "Trafalgar Square", -0.1281, 51.5080,
  "Heathrow Airport", -0.4543, 51.4700,
  "UCL campus", -0.1340, 51.5246
) %>% 
  st_as_sf(coords = c("x", "y"), remove = FALSE, crs = 4326) %>% 
  st_transform(27700) %>% 
  mutate(
    x = unname(st_coordinates(.)[, 1]),
    y = unname(st_coordinates(.)[, 2])
  ) %>% 
  st_set_geometry(NULL)

# plot map
ew_violence_in_grid %>%
  filter(force %in% c("London, City of", "Metropolitan Police")) %>% 
  arrange(desc(crimes)) %>% 
  mutate(
    perc = crimes / sum(crimes),
    cperc = cumsum(perc),
    group = ifelse(cperc < 0.5, "half of all\nviolence", NA_character_)
  ) %>%
  filter(!is.na(group)) %>%
  ggplot() +
  geom_sf(data = london_outline, colour = "white", fill = "white", size = 4) +
  geom_sf(data = london_outline, colour = NA, fill = "white") +
  geom_raster(aes(x = x, y = y, fill = group), na.rm = TRUE, fill = "#F6BE00") +
  geom_sf(data = filter(districts, str_detect(lad19cd, "^E09")), 
          colour = "grey40", fill = NA) +
  ggrepel::geom_label_repel(
    aes(x = x, y = y, label = label), data = london_labels, colour = "grey20", 
    label.size = NA, min.segment.length = 0
  ) +
  geom_point(aes(x = x, y = y), data = london_labels) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = margin(0, 0, 0, 0)
  )  
```



---

class: green-screen, big-picture, animated, fadeIn

```{r camden violence map, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
# create labels for local landmarks
camden_labels <- tribble(
  ~label, ~x, ~y,
  "Camden Town", -0.1426, 51.5390,
  "Covent Garden", -0.1240, 51.5117,
  "Kings Cross", -0.1246, 51.5347,
  "UCL campus", -0.1340, 51.5246
) %>% 
  st_as_sf(coords = c("x", "y"), remove = FALSE, crs = 4326) %>% 
  st_transform(27700) %>% 
  mutate(
    x = unname(st_coordinates(.)[, 1]),
    y = unname(st_coordinates(.)[, 2])
  )

# plot map
ew_violence_in_grid %>%
  filter(district == "Camden") %>% 
  arrange(desc(crimes)) %>% 
  mutate(
    perc = crimes / sum(crimes),
    cperc = cumsum(perc),
    group = ifelse(cperc < 0.5, "half of all\nviolence", NA_character_)
  ) %>%
  filter(!is.na(group)) %>%
  ggplot() +
  geom_sf(data = filter(districts, district == "Camden"), colour = "white", 
          fill = "white", size = 4) +
  geom_sf(data = filter(districts, district == "Camden"), colour = NA, 
          fill = "white") +
  geom_raster(aes(x = x, y = y, fill = group), na.rm = TRUE, fill = "#F6BE00") +
  geom_sf(data = filter(districts, district == "Camden"), colour = "grey40", 
          fill = NA) +
  geom_label(
    aes(x = x, y = y, label = str_wrap(label, 9)), data = camden_labels, 
    colour = "grey20", lineheight = 0.9, label.size = NA
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = margin(0, 0, 0, 0)
  )  
```



---

class: green-screen, animated, fadeIn

<!-- blank slide for green screen fade -->




---

class: big-picture, animated, fadeIn

```{r crime concentration studies map, echo=FALSE}
countries <- rnaturalearth::ne_countries(returnclass = "sf") %>% 
  as("Spatial") %>% 
  maptools::nowrapRecenter() %>% 
  fortify()

concentration_studies <- tribble(
  ~lng, ~lat, ~city, ~perc,
  -75.6972, 45.4215, "Ottawa", 0.017,
  -122.3321, 47.6062, "Seattle", 0.054,
  34.7818, 32.0853, "Tel Aviv", 0.045,
  -74.0060, 40.7128, "New York", 0.05,
  -47.0738, -22.9329, "Campinas", 0.037,
  4.3007, 52.0705, "The Hague", 0.068,
  -93.3563, 45.0941, "Brooklyn Park", 0.02
) %>% 
  mutate(
    label = str_glue("{scales::percent(perc, accuracy = 1)} in {city}")
  )

ggplot() +
  geom_polygon(
    aes(x = long, y = lat), 
    data = countries, 
    colour = NA, 
    fill = "grey25", 
    size = 0.25
  ) +
  geom_label_repel(
    aes(x = lng, y = lat, label = label), 
    data = concentration_studies, 
    fill = alpha("white", 0.67), 
    label.size = NA,
    box.padding = 0.5,
    min.segment.length = 1
  ) +
  geom_point(
    aes(x = lng, y = lat), 
    data = concentration_studies, 
    colour = "red", 
    size = 3
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_map(projection = "ortho", orientation = c(0, 0, 0)) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "black", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = margin(0, 0, 0, 0)
  )
```
